### 2023-12-01 appendix.Rmd
Onde eu descubro que tem multiclonieraridade na aostra com todo mundo e ols não é significativo

---
title: "Data Appendix"
author: "Manoel Galdino"
date: "2023-11-30"
output: 
  bookdown::pdf_document2:
    figsintext : yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)

```

```{r package loads}
library(data.table)
library(stargazer)
library(fixest)
library(here)
library(janitor)
library(lubridate)
library(readxl)
library(tidyverse)
library(gt)
library(ipeadatar)
library(rstan)
library(arm)
library(rstanarm)
library(brms)
library(survey)
library(broom)
library(broom.mixed)
library(MetBrewer)
library(knitr)
library(ggplot2)
library(patchwork)

```

# Appendix

We utilized several data sources in our analysis, with the primary dataset for the quantitative analysis, encompassing the outcome variable, originating from a single source.

The Fundo Nacional de Desenvolvimento da Educação (FNDE), operating under the Ministry of Education in Brazil, is responsible for implementing various policies, including the Proinfância Program. Launched in 2007, Proinfância supports the construction and renovation of schools and nurseries throughout Brazil. Municipalities and states seek funding from FNDE, and in certain cases, FNDE also offers a basic construction project that municipalities have the option to utilize. FNDE maintains a website, SIMEC (an acronym for 'Sistema Integrado de Monitoramento Execução e Controle'), which houses crucial information on all construction projects approved by FNDE. A public version of this data is available at: https://simec.mec.gov.br/painelObras/.

As the dataset undergoes periodic updates with new pieces of information, any alterations may lead to the loss of historical data. To maintain a comprehensive record, Transparência Brasil systematically downloaded spreadsheets from the SIMEC website. These spreadsheets serve as the foundation for the dataset used in our paper.

The initial spreadsheet was collected in May 2017, coinciding with the commencement of the Obra Transparente project. The subsequent spreadsheet was acquired post-project completion, in July 2019. The data set contains information about all construction works approved by FNDE over time. We restricted the analysis only to the kind of projects monitored by Obra Transparente.

## Construction completion

```{r loading data appendix}
load(here("Dados", "obras_inicio_projeto.Rdata"))
load(here("Dados", "obras_fim_seg_fase.Rdata"))

# simec23 <- fread(here("obra_transparente", "Dados", "simec 25-10-23 - simec.csv"), encoding = "UTF-8")
simec23 <- fread(here("Dados", "simec 25-10-23 - simec.csv"), encoding = "UTF-8")

# load(here("obra_transparente", "Dados", "obras_fim_seg_fase.Rdata"))

load(here("Dados", "obras_fim_seg_fase.Rdata"))

df_cidades_ot <- read_excel(here("Dados", "situacao obras_atual.xlsx"), 
                                           sheet = "SIMEC-Mai17", skip = 1) %>%
  clean_names() %>%
  filter(!is.na(id)) %>%
  dplyr::select(id, municipio, uf, situacao2, levant_os_jun_17) %>%
  mutate(indicator_ot = 1)

obras_inicio_projeto_ot <- obras_inicio_projeto %>%
  left_join(df_cidades_ot, by=join_by("municipio", "uf", "id"))

obras_inicio_projeto_ot <- obras_inicio_projeto_ot %>%
  mutate(concluida = grepl("Concluída", situacao))

aux_indicator_ot <- df_cidades_ot %>%
  distinct(municipio, uf) %>%
  mutate(indicator_muni_ot = 1)

obras_inicio_projeto_ot <- obras_inicio_projeto_ot %>%
  mutate(indicator_ot = ifelse(is.na(indicator_ot), 0, indicator_ot)) %>%
  left_join(aux_indicator_ot, by = join_by("municipio", "uf")) %>%
  mutate(indicator_muni_ot = ifelse(is.na(indicator_muni_ot), 0, indicator_muni_ot))

antes_projeto <- obras_inicio_projeto_ot %>%
  dplyr::filter(uf %in% c("SP", "PR", "SP", "SC", "RS", "MG")) %>%
  group_by(indicator_muni_ot, concluida) %>%
  summarise(contagem = n()) %>%
  mutate(total = sum(contagem),
         perc = round(contagem/total, 2))

antes_projeto <- antes_projeto %>%
  dplyr::filter(concluida == TRUE) %>%
  dplyr::select(-c(concluida, contagem, total)) %>%
  ungroup()

obras_fim_seg_fase <- obras_fim_seg_fase %>%
  mutate(concluida = as.numeric(grepl("Concluída", situacao)),
         id = as.numeric(id)) %>%
  left_join(df_cidades_ot, by=join_by("municipio", "uf", "id")) %>%
  mutate(indicator_ot = ifelse(is.na(indicator_ot), 0, indicator_ot)) %>%
  left_join(aux_indicator_ot, by = join_by("municipio", "uf")) %>%
  mutate(indicator_muni_ot = ifelse(is.na(indicator_muni_ot), 0, indicator_muni_ot))

depois_projeto <- obras_fim_seg_fase %>%
  dplyr::filter(uf %in% c("SP", "PR", "SP", "SC", "RS", "MG")) %>%
  group_by(indicator_muni_ot, concluida) %>%
  summarise(contagem = n()) %>%
  mutate(total = sum(contagem),
         perc = round(contagem/total, 2))

depois_projeto <- depois_projeto %>%
  dplyr::filter(concluida == TRUE) %>%
  dplyr::select(-c(concluida, contagem, total)) %>%
  ungroup()

```

Table \@ref(tab:basic-table) illustrates the changes in data before and after the project across the entire country. Notably, there is a significant increase in the completion rate, rising from 37% to 46%, representing an almost 10-percentage-point increase. This change aligns with expectations, considering that the denominator grows at a slower pace than the numerator. The numerator increases with each completion of a construction work, while the denominator expands only with a new funded project.

```{r basic-table}
df_before <- obras_inicio_projeto %>%
  mutate(concluida = as.numeric(grepl("Concluída", situacao))) %>%
  group_by(concluida) %>%
  summarise(num_works = n()) %>%
  mutate(total = sum(num_works)) %>%
  filter(concluida == 1)

df_after <- obras_fim_seg_fase %>%
  mutate(concluida = as.numeric(grepl("Concluída", situacao))) %>%
  group_by(concluida)%>%
  summarise(num_works = n()) %>%
  mutate(total = sum(num_works)) %>%
  filter(concluida == 1)

df1 <- bind_rows(df_before, df_after)

df1 %>%
  rename(date = concluida) %>%
  mutate(date = c("May 2017", "July 2019"),
         perc = scales::percent(round(num_works/total, 2))) %>%
  kable(col.names = c("Date", "Completed construction works", "Total works", "Completion percentage"),
        caption = "Summary of Data Collection Dates and Counts of Completed and Total Construction Works - Brazil")

```

In the main analysis, we restricted the sample to municipalities in the same states where the Obra Transparente project occurred. The reason is that there is perfect multicolinearity between our treatment and seven other control variables (mesured at the municipality level) in the unrestricted sample (log of per capita income, school frequency, log of GDP, share of poor families, illiteracy rate, and an inequality measure).

We present simple descriptive statistics and visualizations to show that the change is similar in our outcome variable considering the restricted and unrestricted sample.

Table \@ref(tab:basic-table-s) presents a situation similar to that of the entire country, albeit with a higher baseline (11 percentage points above the whole country).

```{r basic-table-s}
obras_inicio_projeto_ot <- obras_inicio_projeto %>%
  left_join(df_cidades_ot, by=join_by("municipio", "uf", "id"))

df_before_s <- obras_inicio_projeto %>%
  dplyr::filter(uf %in% c("SP", "PR", "SP", "SC", "RS", "MG")) %>%
  mutate(concluida = as.numeric(grepl("Concluída", situacao))) %>%
  group_by(concluida) %>%
  summarise(num_works = n()) %>%
  mutate(total = sum(num_works)) %>%
  filter(concluida == 1)

df_after_s <- obras_fim_seg_fase %>%
  dplyr::filter(uf %in% c("SP", "PR", "SP", "SC", "RS", "MG")) %>%
  mutate(concluida = as.numeric(grepl("Concluída", situacao))) %>%
  group_by(concluida)%>%
  summarise(num_works = n()) %>%
  mutate(total = sum(num_works)) %>%
  filter(concluida == 1)

df2 <- bind_rows(df_before_s, df_after_s)

df2 %>%
  rename(date = concluida) %>%
  mutate(date = c("May 2017", "July 2019"),
         perc = scales::percent(round(num_works/total, 2))) %>%
  kable(col.names = c("Date", "Completed construction works", "Total works", "Completion percentage"),
        caption = "Summary of Data Collection Dates and Counts of Completed and Total Construction Works in Obra Transparente states")
```

Figure \@ref(fig:plot) presents a comparison of completion rates between restricted (municipalities in the Obra Transparente states operation) and unrestricted samples (the whole country), displayed across different dates (start and end of project).

```{r plot, echo=FALSE, fig.cap="Completion rates between Restricted and Unrestricted samples"}

df1 <- df1 %>%
  rename(date = concluida) %>%
  mutate(date = c("05-17", "07-19"),
         perc = scales::percent(round(num_works/total, 2)))

df2 <- df2 %>%
  rename(date = concluida) %>%
  mutate(date = c("05-17", "07-19"),
         perc = scales::percent(round(num_works/total, 2)))

combined_df <- bind_rows(
  mutate(df1, Sample = "Whole Country"),
  mutate(df2, Sample = "Restriced states")
)

library(ggplot2)

# Assuming combined_df is your dataframe
# You may need to install and load the tidyverse package if you haven't already: install.packages("tidyverse")

# Plotting
ggplot(combined_df, aes(x = date, y = perc, color = date)) +
  geom_point(position = position_dodge(width = 0.5), size = 3) +
  geom_linerange(
    aes(ymin = 0, ymax = perc, color = date),
    position = position_dodge(width = 0.5),
    linewidth = 1.2
  ) +
  facet_wrap(~Sample, ncol = 2, scales = "fixed") +
  labs(
    title = "Comparison of Completion Rates between Restricted and Unrestricted Samples",
    y = "Completion Rate (%)",
    x = NULL
  ) +
  theme_minimal() 
```

## Imbalance

The treatment is just a list of municipalities where the project occurred. To assess imbalance of treatment and control units, we used other metrics in the SIMEC data set (if a construction work is paralyzed or not, it it is in execution or not and percentage completed), since they are related to how hard it is to finish a construction.

Figure \@ref(fig:plot1) shows that there is indeed imbalance among construction works in the restricted sample when compared with the unrestricted sample. Although the rate of change seems to be similar from the beginning and end of project, the baseline is clearly different. It is also worthwhile to note that there are both more construction works in execution and paralyzed in the country as a whole than in the states of Obra Transparente. It is possible because we can also have planned construction (not started), cancelled and completed. Last, but not least, the mean execution rate is higher in the restricted sample, suggesting they were closer to completion in the beggining of the project. These data alone suggest systematic differences between samples.

```{r plot1, echo=FALSE, fig.cap="Completion rates between Restricted and Unrestricted samples"}

imb_before_s <- obras_inicio_projeto %>%
  dplyr::filter(uf %in% c("SP", "PR", "SP", "SC", "RS", "MG")) %>%
  mutate(paralyzed = as.numeric(grepl("Paralisada", situacao)),
         execution = as.numeric(grepl("Execução", situacao)),
         percentage_executed = percentual_de_execucao) %>%
  summarise(num_works = n(),
            num_paralyzed = sum(paralyzed),
            num_execution = sum(execution),
            perc_paralyzed = num_paralyzed/num_works,
            perc_execution = num_execution/num_works,
            mean_execution = mean(percentage_executed)/100) %>%
  mutate(date = "05-17")

imb_before <- obras_inicio_projeto %>%
  mutate(paralyzed = as.numeric(grepl("Paralisada", situacao)),
         execution = as.numeric(grepl("Execução", situacao)),
         percentage_executed = percentual_de_execucao) %>%
  summarise(num_works = n(),
            num_paralyzed = sum(paralyzed),
            num_execution = sum(execution),
            perc_paralyzed = num_paralyzed/num_works,
            perc_execution = num_execution/num_works,
            mean_execution = mean(percentage_executed)/100) %>%
  mutate(date = "05-17")

imb_after_s <- obras_fim_seg_fase %>%
  dplyr::filter(uf %in% c("SP", "PR", "SC", "RS", "MG")) %>%
  mutate(paralyzed = as.numeric(grepl("Paralisada", situacao)),
         execution = as.numeric(grepl("Execução", situacao)),
         percentage_executed = percentual_de_execucao) %>%
  summarise(num_works = n(),
            num_paralyzed = sum(paralyzed),
            num_execution = sum(execution),
            perc_paralyzed = num_paralyzed/num_works,
            perc_execution = num_execution/num_works,
            mean_execution = mean(percentage_executed)/100) %>%
  mutate(date = "07-19")

imb_after <- obras_fim_seg_fase %>%
  mutate(paralyzed = as.numeric(grepl("Paralisada", situacao)),
         execution = as.numeric(grepl("Execução", situacao)),
         percentage_executed = percentual_de_execucao) %>%
  summarise(num_works = n(),
            num_paralyzed = sum(paralyzed),
            num_execution = sum(execution),
            perc_paralyzed = num_paralyzed/num_works,
            perc_execution = num_execution/num_works,
            mean_execution = mean(percentage_executed)/100) %>%
  mutate(date = "07-19")

combined_df <- bind_rows(
  mutate(imb_before_s, Sample = "Restriced states"),
  mutate(imb_before, Sample = "Whole country"),
  mutate(imb_after_s, Sample = "Restriced states"),
  mutate(imb_after, Sample = "Whole country")
  )

df_combined <- combined_df
# Plotting

# Function to create a plot for each variable
create_plot <- function(data, y_variable, title, y_label) {
  ggplot(data, aes(x = date, color = date)) +
    geom_point(aes(y = !!sym(y_variable)), position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.75), size = 3) + scale_y_continuous(labels = scales::percent) +
    geom_linerange(
      aes(ymin = 0, ymax = !!sym(y_variable)),
      position = position_dodge(width = 0.75),
      linewidth = 1.2
    ) +
    facet_wrap(~Sample, scales = "fixed") +
    labs(
      title = title,
      y = y_label,
      x = NULL,
      caption = "Data source: Your Data Source"
    ) +
    theme_minimal() +
    theme(legend.position = "none")
}

# Create individual plots
plot_execution <- create_plot(df_combined, "perc_execution","", "Percentage of construction works in execution")
plot_paralyzed <- create_plot(df_combined, "perc_paralyzed", "", "Percentage of construction works paralyzed")
plot_mean_execution <- create_plot(df_combined, "mean_execution", "", "Mean Execution rate")

# Arrange the plots
combined_plots <- plot_execution + plot_paralyzed + plot_mean_execution

# Print the combined plots
print(combined_plots)
```

In any case, we also ran standard ols regression, adjusting for such variables, as a robustness check.

```{r loading-processing ipea data,  message=FALSE, echo=FALSE, cache=TRUE}
aux_muni <- geobr::lookup_muni(code_muni = "all") %>%
  dplyr::select(-c(name_state, name_intermediate))

ipea_idhm <- ipeadatar::ipeadata("ADH_IDHM") %>%
  dplyr::filter(date > "2001-01-1" & as.character(uname) == "Municipality") %>%
  rename(idhm = value)

ipea_pea18 <- ipeadatar::ipeadata("ADH_PEA18M")  %>%
  dplyr::filter(date > "2001-01-1" & as.character(uname) == "Municipality") %>%
  rename(pea18 = value)

ipea_ext_pobres <- ipeadatar::ipeadata("ADH_PIND")  %>%
  dplyr::filter(date > "2001-01-1" & as.character(uname) == "Municipality") %>%
  rename(ext_pobres = value)

ipea_ext_pobres_neg <- ipeadatar::ipeadata("ADH_PIND_NEG")  %>%
  dplyr::filter(date > "2001-01-1" & as.character(uname) == "Municipality") %>%
  rename(ext_pobres_negros = value)

ipea_rpc <- ipeadatar::ipeadata("ADH_RDPC")  %>%
  dplyr::filter(date > "2001-01-1" & as.character(uname) == "Municipality") %>%
  rename(rpc = value)

ipea_rpc_mulheres <- ipeadatar::ipeadata("ADH_RDPC_MUL")  %>%
  dplyr::filter(date > "2001-01-1" & as.character(uname) == "Municipality") %>%
  rename(rpc_mulheres = value)

ipea_rpc_negros <- ipeadatar::ipeadata("ADH_RDPC_NEG")  %>%
  dplyr::filter(date > "2001-01-1" & as.character(uname) == "Municipality") %>%
  rename(rpc_negros = value)

ipea_ext_prop_pobres <- ipeadatar::ipeadata("ADH_PMPOB")  %>%
  dplyr::filter(date > "2001-01-1" & as.character(uname) == "Municipality") %>%
  rename(ext_prop_pobres = value)

ipea_10p_ricos_desigualdade <- ipeadatar::ipeadata("ADH_PREN10RICOS")  %>%
  dplyr::filter(date > "2001-01-1" & as.character(uname) == "Municipality") %>%
  rename(p10_ricos_desigualdade = value)

ipea_freq_escola_15_17 <- ipeadatar::ipeadata("ADH_T_FREQ15A17")  %>%
  dplyr::filter(date > "2001-01-1" & as.character(uname) == "Municipality") %>%
  rename(freq_escola_15_1 = value)

ipea_analfabetismo <- ipeadatar::ipeadata("ADH_T_ANALF15M")  %>%
  dplyr::filter(date > "2001-01-1" & as.character(uname) == "Municipality") %>%
  rename(analfabetismo = value)

pop2010 <- ribge::habitantes2010 %>%
  mutate(tcode = as.integer(paste0(codigo_uf, codigo_munic)))

ipea_final <- ipea_idhm %>%
  inner_join(aux_muni, by = join_by(tcode == code_muni)) %>%
  inner_join(ipea_pea18, by = "tcode") %>%
  inner_join(ipea_ext_pobres, by = "tcode") %>%
  inner_join(ipea_rpc, by = "tcode") %>%
  inner_join(ipea_ext_prop_pobres, by = "tcode") %>%
  inner_join(ipea_10p_ricos_desigualdade, by = "tcode") %>%
  inner_join(ipea_freq_escola_15_17, by = "tcode") %>%
  inner_join(ipea_analfabetismo, by = "tcode") %>%
  inner_join(pop2010, by = "tcode") %>%
  dplyr::select(-contains("date")) %>%
  dplyr::select(-contains("uname")) %>%
  dplyr::select(-starts_with("code"))
```

```{r processing reg, message=FALSE, echo=FALSE}
aux_indicator_ot <- df_cidades_ot %>%
  distinct(municipio, uf) %>%
  mutate(indicator_muni_ot = 1)

obras_inicio_projeto_ot <- obras_inicio_projeto_ot %>%
  left_join(aux_indicator_ot, by = join_by("municipio", "uf")) %>%
  mutate(indicator_muni_ot = ifelse(is.na(indicator_muni_ot), 0, indicator_muni_ot))

aux_muni1 <- aux_muni %>%
  mutate(name_muni1 = tolower(iconv(name_muni, "UTF-8",  "ASCII//TRANSLIT")))

df_balancing_aux <- obras_inicio_projeto_ot %>%
  mutate(concluida = as.numeric(grepl("Concluída", situacao)))  %>%
  mutate(municipio1 = tolower(iconv(municipio, "UTF-8",  "ASCII//TRANSLIT"))) %>%
  inner_join(aux_muni1, by = join_by(municipio1  == name_muni1, uf == abbrev_state)) %>%
  inner_join(ipea_final, by = join_by(code_muni == tcode)) %>%
  mutate(perc_pea18 =pea18/populacao,
         log_pop = log(populacao),
         log_renda = log(rpc*populacao),
         log_rpc = log(rpc),
         tipo_do_projeto = ifelse(grepl("Bloco Estrutural" ,tipo_do_projeto),
                                  "Projeto Tipo B ou C - Bloco Estrutural", tipo_do_projeto)) %>%
  mutate(concluida_num = as.numeric(concluida))

df_balancing_aux1 <- df_balancing_aux  %>%
  dplyr::filter(uf.x %in% c("SP", "MG", "RS", "PR", "SC"))

pre_dados <- df_balancing_aux %>%
  dplyr::select(id, concluida) %>%
  rename(concluida_pre = concluida) %>%
  mutate(id = as.numeric(id))

obras_fim_seg_fase <- obras_fim_seg_fase %>%
  mutate(concluida = as.numeric(grepl("Concluída", situacao)),
         id = as.numeric(id))

data_pos <- obras_fim_seg_fase %>%
  mutate(municipio1 = tolower(iconv(municipio, "UTF-8",  "ASCII//TRANSLIT"))) %>%
  inner_join(aux_muni1, by = join_by(municipio1  == name_muni1, uf == abbrev_state)) %>%
  inner_join(ipea_final, by = join_by(code_muni == tcode)) %>%
  mutate(perc_pea18 =pea18/populacao,
         log_pop = log(populacao),
         log_renda = log(rpc*populacao),
         log_rpc = log(rpc)) %>%
  inner_join(pre_dados, by = "id") %>%
  dplyr::filter(concluida_pre == FALSE) %>%
  mutate(concluida_num = as.numeric(concluida))

df_balancing_aux1 <- df_balancing_aux1 %>%
  dplyr::select(concluida_num , indicator_muni_ot, code_muni, log_rpc, ext_pobres, freq_escola_15_1,analfabetismo,
                log_renda, p10_ricos_desigualdade, percentual_de_execucao, municipio) %>%
  mutate(code_muni = as.factor(code_muni))

te_spec_nr <- formula(concluida_num ~ log_rpc  + ext_pobres + freq_escola_15_1 + analfabetismo +
                     log_renda  +  p10_ricos_desigualdade + percentual_de_execucao + indicator_muni_ot + (1 | code_muni) + .)

```

Tables below presents results from standard, frequentist regression (ols, fixed effects, both with clustered standard errors) and Bayesian regression (linear model with varying intercept).

```{r fitting, echo=FALSE, message=FALSE, results="asis"}

# Simple linear regression
fit_ols <- feols(concluida_num ~ indicator_muni_ot + log_rpc +
                   ext_pobres + freq_escola_15_1 + analfabetismo + log_renda + p10_ricos_desigualdade + percentual_de_execucao, data = df_balancing_aux1, cluster = "code_muni" )

fit_fixed <- feols(concluida_num ~ indicator_muni_ot + log_rpc + ext_pobres + freq_escola_15_1 + analfabetismo + log_renda + p10_ricos_desigualdade + percentual_de_execucao |code_muni, data = df_balancing_aux1, cluster = "code_muni")

etable(list(fit_ols, fit_fixed), tex=TRUE)

set.seed(120)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
# system.time(reg_ps_pre <- brm(te_spec_nr, data=df_balancing_aux1, iter = 100, cores = 4, chains = 4))
# reg_ps_pre <- stan_lmer(te_spec_nr, data=df_balancing_aux1, iter = 100)

## After
# set.seed(120)
# reg_ps_pos <- stan_lmer(te_spec_nr, data=df_balancing_aux_post_se[matches_pos$match.ind,], iter = 4000, refresh=0)

```

```{r processing bayes fit, echo=FALSE,  message=FALSE}
# posterior draws

##
# Before
##

# b1_pre_sims <- as.matrix(reg_ps_pre, 
#                      pars = "indicator_muni_ot")

# Posterior mean and SD of treatment
# b1_mean_pre <- mean(b1_pre_sims) # posterior mean
# b1_sd_pre <- sd(b1_pre_sims)       # posterior SD
#

# Posterior median and 95% credible interval
# b1_quant_pre <- quantile(b1_pre_sims,   probs = c(0.025, 0.50, 0.975))
# 
# b1_quant_pre <- data.frame(t(b1_quant_pre))
# names(b1_quant_pre) <- c("Q2.5", "Q50", "Q97.5")

# Combine summary statistics of posterior simulation draws
# b1_df_pre_cred <- data.frame(variables = "ATT before project",
#                              post_mean = b1_mean_pre,
#                              post_sd = b1_sd_pre,
#                              b1_quant_pre[2], 
#                              b1_quant_pre[1],
#                              b1_quant_pre[3])
#

##
# After
##
# b1_sims_after <- as.matrix(reg_ps_pos, 
#                      pars = "indicator_muni_ot")
#   
# Compute mean, SD, median, and 95% credible interval of treatment

# Posterior mean and SD of treatment
# b1_mean_after <- mean(b1_sims_after) # posterior mean
# b1_sd_after <- sd(b1_sims_after)       # posterior SD
# 
# # Posterior median and 95% credible interval
# b1_quant_after <- quantile(b1_sims_after,   probs = c(0.025, 0.50, 0.975))
# 
# b1_quant_after <- data.frame(t(b1_quant_after))
# names(b1_quant_after) <- c("Q2.5", "Q50", "Q97.5")

# Combine summary statistics of posterior simulation draws
# b1_df_after_cred <- data.frame(variables = "ATT After project",
#                              post_mean = b1_mean_after,
#                              post_sd = b1_sd_after,
#                              b1_quant_after[2], 
#                              b1_quant_after[1],
#                              b1_quant_after[3])
# b1_df <- bind_rows(b1_df_pre_cred, b1_df_after_cred)
```

```{r table2, echo=FALSE,  message=FALSE, results='asis'}

# b1_df %>% gt() %>%
#   tab_header(
#     title = "Effect of local OSB social monitoring on construction delivery rates",
#     subtitle = "Posterior probabilities") %>%
#   cols_label(
#     variables = "Variables",
#     post_mean = "Mean",
#     post_sd = "Sd",
#     Q50 = "Median",
#     Q2.5 = "2.5% CI",
#     Q97.5 = "97.5% CI"
#   ) %>%
#   fmt_number(
#     columns = c(post_mean, post_sd, Q50, Q2.5, Q97.5 ),
#     decimals = 3,
#     use_seps = FALSE
#   ) %>%
#   tab_source_note(
#     source_note = "452 observations for the pre-treatment period and 320 observations for the post-treatment period, resulting from the matching procedure, and encompassing municipalities in the same states as municipalities of Obra Transparente"
#   ) %>%
#   tab_options(., container.width = 500, container.height =     500)

```